Hydra Battles and Cie

### 2021-02-15 doc/chapter-powers.tex
moving sub-library MoreAck to theories/ordinals

@@

\begin{todo}
Put links when done.
\end{todo}

@@

@@

@@

\section{Addition chains}
An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
intermediate steps that lead to the evaluation of some $x^n$, under the 
assumption that each of these steps is a computation of  a power $x^i$, with 
$i<n$.

@@

The following \emph{monadic} 
notation makes terms of type \texttt{computation} look like
expressions of a small programming language dedicated to sequences of mutiplications.
Please look at \emph{CPDT}~\cite{chlipalacpdt2011} for more details on monadic notations in \coq.

@@

@@

\section{Euclidean Chains}
\label{euclide-sect}

In this section, we present an efficient chain generator. The chains built by this generator
are never longer than the chains built by the binary generator. Moreover, for an 
infinite number of exponents, the chains it builds are strictly shorter than the chain
returned by \texttt{binary\_chain}. 
Euclidean chains are based on the following idea: 
\begin{quote}
For generating a chain that computes $x^n$, one may choose some natural number
$0<p<n$, and build a chain that computes first $x^p$ \textbf{then} uses this value
for computing $x^n$. 
\end{quote}

>>>>

\emph{The \texttt{Monoid} type class is defined 
page~\pageref{sect:monoid-def}.}

@@

@@

\label{sect:monoid-def}

\index{additions}{Type classes!Monoid}

@@

\index{additions}{Type classes!EMonoid}

@@

\section{Addition chains}
\index{maths}{Addition chains}
An \emph{addition chain} (In short : \emph{a chain})~\cite{brauer1939} is a representation of a sequence of
intermediate steps that lead to the evaluation of  $x^n$, under the 
assumption that each of these steps is a computation of  a power $x^i$, with 
$i<n$.

@@

\index{additions}{Types!computation}

The following \emph{monadic} 
notation makes terms of type \texttt{computation} look like
expressions of a small programming language dedicated to sequences of mutiplications.
Please look at \emph{CPDT}~\cite{chlipalacpdt2011} for more details on monadic notations in \coq.
\label{monadic-mult}

@@

\index{additions}{Types!chain@chain (addition chains)}

@@

\section{Euclidean Chains}
\label{euclide-sect}
\index{maths}{Euclidean addition chains}
In this section, we present an efficient chain generator. The chains built by this generator
are never longer than the chains built by the binary generator. Moreover, for an 
infinite number of exponents, the chains it builds are strictly shorter than the chain
returned by \texttt{binary\_chain}. 
Euclidean chains are based on the following idea: 
\begin{quote}
For generating a chain that computes $x^n$, one may choose some natural number
$0<p<n$, and build a chain that computes first $x^p$ \textbf{then} uses this value
for computing $x^n$. 
\end{quote}

### 2021-02-15 doc/chapter-primrec.tex
creation of a multiple index
work on the index
moving sub-library MoreAck to theories/ordinals

O'Connor's library on Gödel's incompleteness theorems contains a little more 
than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
are stored in a sub-directory \texttt{Ackermann/MoreAck}.

\section{Basic Definitions}
\index{Maths!Primitive recursive functions}

@@

\begin{Coqsrc}
Fixpoint naryFunc (n : nat) : Set :=
  match n with
  | O => nat
  | S n => nat -> naryFunc n
  end.
\end{Coqsrc}

@@

\vspace{4pt}
\noindent
\emph{From\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
\begin{Coqsrc}
Require Import primRec.
Import extEqualNat.

@@

\index{Coq!Dependent types}
\index{Coq!Dependently typed functions}
\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}
\begin{Coqsrc}
Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
  match n with
    0 => fun a b => a = b
  | S p => fun a b => forall c, extEqual p (a c) (b c)
  end.
\end{Coqsrc}

@@

\index{Coq!Dependent types}
\index{Coq!Mutually inductive types}

\label{def:Primrec}
\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
\begin{Coqsrc}
Inductive PrimRec : nat -> Set :=
  | succFunc : PrimRec 1
  | zeroFunc : PrimRec 0
  | projFunc : forall n m : nat, m < n -> PrimRec n
  | composeFunc :
      forall (n m : nat) (g : PrimRecs n m) (h : PrimRec m), PrimRec n
  | primRecFunc :
      forall (n : nat) (g : PrimRec n) (h : PrimRec (S (S n))), 
      PrimRec (S n)
with PrimRecs : nat -> nat -> Set :=
  | PRnil : forall n : nat, PrimRecs n 0
  | PRcons : forall n m : nat, PrimRec n -> PrimRecs n m -> 
                   PrimRecs n (S m).
\end{Coqsrc}

@@

\index{Coq!Dependent pattern matching}
Both functions are mutually defined through dependent pattern matching. We advise the readers who 
would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look also at the helper functions in
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.

@@

@@

The constructors \texttt{zeroFunc}, \texttt{succFunc},  and \texttt{projFunc} of type
\texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
Although  the following lemmas are alreday proven in 
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
we wrote alternate proofs in 
\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}%
{Ackermann.MoreAck.PrimRecExamples.v}, in order to illustrate the main proof patterns.

@@

Projections are proved primitive recursive, case by case (many examples in 
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v}).
\emph{Please notice again that the name of the projection follows the mathematical tradition, 
whilst the arguments of  \texttt{projFunc} use another convention (\emph{cf} remark~\vref{projFunc-order-of-args}).}

@@

@@

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}}.
\begin{Coqsrc}
Lemma  const1_NIsPR n : isPR 0 n. 
Proof.
  induction n.
  - apply zeroIsPR.
  - destruct IHn as [x Hx].
   exists (composeFunc 0 1 (PRcons 0 0 x (PRnil 0)) succFunc). 
   cbn in *; intros; now rewrite Hx.
Qed.
\end{Coqsrc}

@@

The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}. We present in 
\href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}
a commented version of this proof,

@@

\begin{todo}
Comment more examples from   \href{../theories/html/hydras.Ackermann.MoreAck.PrimRecExamples.html}{Ackermann.MoreAck.PrimRecExamples}.
\end{todo}

\begin{exercise}
There is a lot of lemmas similar to \texttt{filter010IsPR} in the \texttt{primRec} library, useful to control the arity of functions.
Thus, the reader may look at them, and invent simple examples of application for each one.
\end{exercise}

\begin{exercise}
Multiplication of natural number is already proven in the \texttt{primRec} library. Write a proof of your own, then compare to the library's version.
\end{exercise}

@@

\begin{exercise}
Consider the following functions:

@@

\index{Exercises}

@@

\index{Exercises}
\begin{exercise}
Show that the function \texttt{min: naryFunc\,2} is primitive
recursive.

@@

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.

\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.Ack.html}{Ackermann.MoreAck.Ack}}.
\index{Maths!Ackermann function}

@@

\emph{From \href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}}.
\begin{Coqsrc}
Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
\end{Coqsrc}

@@

\begin{Coqsrc}
Scheme PrimRec_PrimRecs_ind := Induction for PrimRec
  Sort Prop
  with PrimRecs_PrimRec_ind := Induction for PrimRecs 
  Sort Prop.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint max_v {n:nat} : forall (v: Vector.t nat n) , nat :=
  match n as n0 return (Vector.t nat n0 -> nat)
  with
    0 => fun v => 0
  | S p => fun (v : Vector.t nat (S p)) =>
             (max (Vector.hd v) (max_v  (Vector.tl v)))
  end.

@@

In Library\href{../theories/html/hydras.Ackermann.MoreAck.AckNotPR.html}{Ackermann.MoreAck.AckNotPR}, we introduce a lighter notation.

\begin{Coqsrc}
(**  uncurried apply:

@@

\begin{Coqsrc}
Definition majorized {n} (f: naryFunc n) (A: naryFunc 2) : Prop :=
  exists (q:nat), forall (v: t nat n),
      v_apply f v <= A q  (max_v v).

@@

>>>>

O'Connor's library on Gödel's incompleteness theorems contains a little more 
than 45K lines of scripts. The part dedicated to primitive recursive functions and Peano arithmetics is 32K lines long and is originally structured in 38 modules.
Thus, we propose a slow exploration of this library, through examples and exercises. Our additions to the original library 
are stored in the directory \texttt{theories/ordinals/MoreAck}.

\section{Basic Definitions}
\index{maths}{Primitive recursive functions}

@@

\index{primrec}{Types!naryFunc}
\begin{Coqsrc}
Fixpoint naryFunc (n : nat) : Set :=
  match n with
  | O => nat
  | S n => nat -> naryFunc n
  end.
\end{Coqsrc}

@@

\vspace{4pt}
\noindent
\emph{From\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
\begin{Coqsrc}
Require Import primRec.
Import extEqualNat.

@@

\index{coq}{Dependent types}
\index{coq}{Dependently typed functions}
\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.extEqualNat.html}{Ackermann.extEqualNat}}

\index{primrec}{Predicates!extEqual}

\begin{Coqsrc}
Fixpoint  extEqual (n : nat) : forall  (a b : naryFunc n), Prop :=
  match n with
    0 => fun a b => a = b
  | S p => fun a b => forall c, extEqual p (a c) (b c)
  end.
\end{Coqsrc}

@@

\index{coq}{Dependent types}
\index{coq}{Mutually inductive types}

\index{primrec}{Types!PrimRec}
\index{primrec}{Types!PrimRecs}
\label{def:Primrec}
\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.}
\begin{Coqsrc}
Inductive PrimRec : nat -> Set :=
  | succFunc : PrimRec 1
  | zeroFunc : PrimRec 0
  | projFunc : forall n m : nat, m < n -> PrimRec n
  | composeFunc :
      forall (n m : nat) (g : PrimRecs n m) (h : PrimRec m), PrimRec n
  | primRecFunc :
      forall (n : nat) (g : PrimRec n) (h : PrimRec (S (S n))), 
      PrimRec (S n)
with PrimRecs : nat -> nat -> Set :=
  | PRnil : forall n : nat, PrimRecs n 0
  | PRcons : forall n m : nat, PrimRec n -> PrimRecs n m -> 
                   PrimRecs n (S m).
\end{Coqsrc}

@@

\index{primrec}{Functions!evalPrimRec}
\index{primrec}{Functions!evalPrimRecs}

\index{coq}{Dependent pattern matching}
Both functions are mutually defined through dependent pattern matching. We advise the readers who 
would feel uneasy with dependent types to consult Adam Chlipala's \emph{cpdt}  book~\cite{chlipalacpdt2011}. We leave it to the reader  to look also at the helper functions in
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec}.

@@

\index{primrec}{Predicates!isPR}
\index{coq}{Extensionnaly equal functions}

@@

The constructors \texttt{zeroFunc}, \texttt{succFunc},  and \texttt{projFunc} of type
\texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
Although  the following lemmas are alreday proven in 
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
we wrote alternate proofs in 
\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}%
{Ackermann.MoreAck.PrimRecExamples.v}, in order to illustrate the main proof patterns.

@@

Projections are proved primitive recursive, case by case (many examples in 
\href{../theories/html/hydras.primRec.html}{Ackermann.primRec.v}).
\emph{Please notice again that the name of the projection follows the mathematical tradition, 
whilst the arguments of  \texttt{projFunc} use another convention (\emph{cf} remark~\vref{projFunc-order-of-args}).}

@@

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}.}

@@

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}}.
\begin{Coqsrc}
Lemma  const1_NIsPR n : isPR 0 n. 
Proof.
  induction n.
  - apply zeroIsPR.
  - destruct IHn as [x Hx].
   exists (composeFunc 0 1 (PRcons 0 0 x (PRnil 0)) succFunc). 
   cbn in *; intros; now rewrite Hx.
Qed.
\end{Coqsrc}

@@

The lemma \texttt{plusIsPR} is already proven in \href{../theories/html/hydras.primRec.html}{Ackermann.primRec}. We present in 
\href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}
a commented version of this proof,

@@

\begin{todo}
Comment more examples from   \href{../theories/html/hydras.MoreAck.PrimRecExamples.html}{MoreAck.PrimRecExamples}.
\end{todo}

\index{primrec}{Exercises}
\begin{exercise}
There is a lot of lemmas similar to \texttt{filter010IsPR} in the \texttt{primRec} library, useful to control the arity of functions.
Thus, the reader may look at them, and invent simple examples of application for each one.
\end{exercise}

\index{primrec}{Exercises}
\begin{exercise}
Multiplication of natural number is already proven in the \texttt{primRec} library. Write a proof of your own, then compare to the library's version.
\end{exercise}

@@

\index{primrec}{Exercises}
\begin{exercise}
Consider the following functions:

@@

\index{primrec}{Exercises}

@@

\index{primrec}{Exercises}
\begin{exercise}
Show that the function \texttt{min: naryFunc\,2} is primitive
recursive.

@@

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.
\index{hydras}{Library Prelude!iterate}
\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.
\index{maths}{Ackermann function}
\index{primrec}{Ackermann function}

@@

\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
\begin{Coqsrc}
Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
\end{Coqsrc}

@@

\index{coq}{Commands!Scheme}
\begin{Coqsrc}
Scheme PrimRec_PrimRecs_ind := Induction for PrimRec
  Sort Prop
  with PrimRecs_PrimRec_ind := Induction for PrimRecs 
  Sort Prop.
\end{Coqsrc}

@@

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
Fixpoint max_v {n:nat} : forall (v: Vector.t nat n) , nat :=
  match n as n0 return (Vector.t nat n0 -> nat)
  with
    0 => fun v => 0
  | S p => fun (v : Vector.t nat (S p)) =>
             (max (Vector.hd v) (max_v  (Vector.tl v)))
  end.

@@

In Library\href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}, we introduce a lighter notation.

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
(**  uncurried apply:

@@

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
Definition majorized {n} (f: naryFunc n) (A: naryFunc 2) : Prop :=
  exists (q:nat), forall (v: t nat n),
      v_apply f v <= A q  (max_v v).

@@

\index{coq}{Mutual induction}

### 2021-02-15 doc/hydras.tex
moving sub-library MoreAck to theories/ordinals

\SetWatermarkLightness{0.8}
\definecolor{lightorange}{rgb}{1,0.7,0.5}
\SetWatermarkScale{0.35}
\SetWatermarkText{Work in Progress}
\SetWatermarkColor{lightorange}

\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, \\with contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury\\ \emph{The formalization of primitive recursive functions was originally authored by Russel O'Connor}}
\date{\today}
\title{Hydra Battles and Cie}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\maketitle

@@

\include{part-hydras}
\include{chapter-primrec}
\part{Some Funny Algorithms}

@@

\section{Hydras and Ordinals}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Main predicates}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
\hline
\end{tabular}
\begin{tablenotes}
  \item[1] This order is total, but not well-founded, because of not well formed terms.
\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
\item[3] This order is total \emph{and} well-founded.
\end{tablenotes}
\end{threeparttable}
\end{table}

\vspace{4pt}
\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Infix notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\

\hline
\end{tabular}

\begin{tablenotes}
 \item [1]
Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Other notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
\pageref{sect:notation-FS}\\ 
omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

\hline
\end{tabular}

\begin{tablenotes}
  \item [2] Note that there exist also various coercions from \texttt{nat} to types of ordinal. Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
\end{tablenotes}
  \end{threeparttable}

@@

\section{Addition Chains}

\begin{todo}
\end{todo}

>>>>

\SetWatermarkLightness{0.8}
\definecolor{lightorange}{rgb}{1,0.9,0.9}
\SetWatermarkScale{0.35}
\SetWatermarkText{Work in Progress}
\SetWatermarkColor{lightorange}

\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, \thanks{With contributions by Yves Bertot, \'Evelyne Contejean,  Florian Hatat and Pascal Manoury. \emph{The formalization of primitive recursive functions was originally authored by 
Russel O'Connor\cite{OConnor05}.}}}
\date{\today}
\title{Hydra Battles and Cie}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\maketitle

@@

\include{part-hydras}
\include{chapter-primrec}
\part{Some  Certified Algorithms}

@@

\section{Main notations}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Ordinals and ordinal notations}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\

F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
\pageref{sect:notation-FS}\\ 
omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

\hline
\end{tabular}
\begin{tablenotes}
  \item[1] This order is total, but not well-founded, because of not well formed terms.
\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
\item[3] This order is total \emph{and} well-founded.
\item [4]
Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
  \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal. Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
\end{tablenotes}
\end{threeparttable}
\end{table}

\vspace{4pt}

\begin{table}[H]

  \begin{threeparttable}
    \caption{hydra battles}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
\hline
\end{tabular}

%\begin{tablenotes}
%\end{tablenotes}

  \end{threeparttable}

@@

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Addition chains}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
\hline
\end{tabular}

%\begin{tablenotes}
%\end{tablenotes}

  \end{threeparttable}
